---
title: "EDA"
author: "Shafeek Saleem"
date: "11/21/2020"
output:
  pdf_document: default
  html_document: default
---

## Setup

### Load packages

```{r load-packages, message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(xgboost)
library(MLmetrics)
library(lubridate)
library(corrplot)
library(truncnorm)
library(gridExtra)
```

```{r}
setwd('C:\\Users\\octavei\\OneDrive - John Keells Holdings PLC\\Desktop\\MyFiles\\UoC Mini Hackathon')
```

### Load data

```{r load-data}
train <- read.csv("train_data.csv",na.strings = "")
test <- read.csv("test_data.csv",na.strings = "")
```

### Change order of variables 
```{r}
train <- train %>% select(colnames(test),"app_status")
```

* * *

## Explonatory Data Analysis

### Feature description

* * *

The features used in the structure are:

   NURSERY            Evaluation of applications for nursery schools
   . EMPLOY           Employment of parents and child's nursery
   . . parents        Parents' occupation
   . . has_nurs       Child's nursery
   . STRUCT_FINAN     Family structure and financial standings
   . . STRUCTURE      Family structure
   . . . form         Form of the family
   . . . children     Number of children
   . . housing        Housing conditions
   . . finance        Financial standing of the family
   . SOC_HEALTH       Social and health picture of the family
   . . social         Social conditions
   . . health         Health conditions

* * *

### Initial analysis
```{r}
str(train)
```
**Decisions -**

1. `app_status` should be Factor variables.
2. `ID` should be removed.


```{r}
train <- train %>% select(-ID)
test <- test %>% select(-ID)
train$app_status <- as.factor(train$app_status)
train$has_nurs = factor(train$has_nurs, levels = c("proper", "less_proper", "improper", "critical", "very_crit"))
train$housing = factor(train$housing, levels = c("convenient", "less_conv", "critical"))
train$social = factor(train$social, levels = c("nonprob", "slightly_prob", "problematic"))
train$form = factor(train$form, levels = c("complete", "completed", "incomplete", "foster"))
```

* * *

### Handling missing/ null values

```{r}
train %>% summarise(Feature = colnames(train),"Total MissingValues" = colSums(is.na(train)),"MissingValues(%)" = round(colSums(is.na(train))/n()*100,2))
test %>% summarise(Feature = colnames(test),"Total MissingValues" = colSums(is.na(test)), "MissingValues(%)" = round(colSums(is.na(test))/n()*100,2))
```

**Observations and Decisions -** 

1. No missing or null values found in both train and test datasets.

* * *

### Target class/Y distribution

Univariate analysis of Target variable : Survived
```{r}
train %>% group_by(app_status) %>% summarise(perc = n()*100/nrow(train))
train %>% ggplot(aes(x=as.factor(app_status)))+geom_bar(fill = "#56B4E9")+labs(x='app_status',y='Count',title = "Target Class Distribution")
```

**Note -**

 Target class distribution is slightly imbalanced - Train dataset contains 64.5% of class 0 and 35.5% of class 1. One good solution would be up sampling which can cuase over fitting. Another solution would be to use weight parameters with algorithms like xgboost. 

* * *

* * *

## Multivariate analysis - Features vs Target
* * *

## 1. Employment
* * *
### 1.1. Employment of parents

```{r}
train %>%
  ggplot(aes(x=parents, fill=app_status))+geom_bar(position="fill")+labs(x='Parents Occupation',y='Percentage')

train %>% 
  group_by(parents) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. Applications with, parent's employment as usual, have ~30% more chance to get accepted than parents with great_pret employment. 

* * *

### 1.2. Child's nursery

```{r}
train %>%
  ggplot(aes(x=has_nurs, fill=app_status))+geom_bar(position="fill")+coord_flip() +labs(x="Child's nursery",y='Percentage')

train %>% 
  group_by(has_nurs) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. The probability of an application getting accepted is increasing with the child's nursery from very cretical to proper (~8% to ~58%). 

* * *

## 2. Family structure and Financial standings
* * *
### 2.1. Family structure

```{r}
plot1 <- train %>%
  ggplot(aes(x=form, fill=app_status))+
  geom_bar(position="fill")+
  labs(x='Form of the family',y='Percentage')+
  scale_fill_manual(values = c("#FFDB6D", "#C4961A"))+
  coord_flip() 

plot2 <- train %>%
  ggplot(aes(x=children, fill=app_status))+
  geom_bar(position="fill")+
  labs(x='Number of children',y='Percentage')+
  scale_fill_manual(values = c("#56B4E9", "#009E73"))

grid.arrange(plot1, plot2, ncol=2)

train %>% 
  group_by(form) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')

train %>% 
  group_by(children) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** class  distribution of both classes are balanced.

**Observations -**

1. Applications with, complete family form have ~7% more chance to be accepted than foster families. The order for probabilty of acceptance is foster < incomplete < completed < complete.

2. The probability of an application getting accepted is decreasing with the number of children in a family (~41% to ~32%). 

* * *

### 2.2. Housing conditions

```{r}
train %>%
  ggplot(aes(x=housing, fill=app_status))+
  geom_bar(position="fill")+coord_flip()+
  labs(x="Housing conditions",y='Percentage')+
  scale_fill_manual(values = c("#56B4E9", "#009E73"))

train %>% 
  group_by(housing) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. Applications with, convenient housing conditions have ~13%% more chance to be accepted than applications with critical house conditions.


### 2.3. Financial standing of the family

```{r}
train %>%
  ggplot(aes(x=finance, fill=app_status))+
  geom_bar(position="fill")+
  labs(x="Financial standing of the family",y='Percentage')+
  scale_fill_manual(values = c("#56B4E9", "#009E72"))

train %>% 
  group_by(finance) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. Applications with, convenient financial standings have ~5% more chance to be accepted than applications with inconvenient financial standings.
* * *

## 3. Social and health picture of the family
* * *

### 3.1. Social conditions

```{r}
train %>%
  ggplot(aes(x=social, fill=app_status))+
  geom_bar(position="fill")+
  labs(x="Social conditions",y='Percentage')+
  scale_fill_manual(values = c("yellow4", "yellow3"))

train %>% 
  group_by(social) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. Applications with non-problemetic or slightly-problemetic social conditions have almost equal chance of getting accepted.
2. But applications with problemetic social conditions have ~10% lesser chance of getting accepted than the other two categories.
* * *

### 3.2. Health conditions

```{r}
train %>%
  ggplot(aes(x=health, fill=app_status))+
  geom_bar(position="fill")+
  labs(x="Health conditions",y='Percentage')+
  scale_fill_manual(values = c("#56B4E9", "#009E73"))

train %>% 
  group_by(health) %>% 
  summarise(Count = n(), "App status Rate ( % )" = sum(app_status==1)/n()*100, .groups = 'drop')
```

**Note -** Feature class  distribution is balanced.

**Observations -**

1. Applications with not-recommended health conditions have 100% chance of getting rejected. Health conditions play a major role in selection of applications.
2. Applications with recommended health conditions have ~64% chance of getting accepted.
3. Applications with priority health conditions have ~43% chance of getting accepted.

* * *

